### 2019-09-03 paper/HOLMultivariate.tex

\newcommand{\HOLMultivariateDate}{03 Settembre 2019}
\newcommand{\HOLMultivariateTime}{10:42}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXequationXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs}) \HOLBoundVar{Es}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXequationXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXequationXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXsolutionXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{R} \HOLBoundVar{Ps}.
       \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{R} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{e}) (\HOLConst{set} \HOLBoundVar{Xs})) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{LIST_REL} \HOLBoundVar{R} \HOLBoundVar{Ps} (\HOLConst{MAP} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps})) \HOLBoundVar{Es})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXsolutionXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXsolutionXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXSUBSTXXdef}
\HOLTokenTurnstile{} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLConst{nil} \HOLSymConst{=} \HOLConst{nil}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{=} \HOLBoundVar{u}\HOLSymConst{..}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{=}
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{=}
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{=} \HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E} \HOLBoundVar{rf}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{=} \HOLConst{relab} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}) \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{var} \HOLBoundVar{X}) \HOLSymConst{=}
        \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLBoundVar{fm} \HOLConst{'} \HOLBoundVar{X} \HOLKeyword{else} \HOLConst{var} \HOLBoundVar{X}) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X} \HOLBoundVar{E}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} \HOLBoundVar{E}) \HOLSymConst{=}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} (\HOLConst{CCS_SUBST} (\HOLBoundVar{fm} \HOLSymConst{\HOLTokenBackslash{}\HOLTokenBackslash{}} \HOLBoundVar{X}) \HOLBoundVar{E})
       \HOLKeyword{else} \HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXSUBSTXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXSUBSTXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionscontextXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}.
       \HOLConst{context} \HOLBoundVar{Xs} \HOLSymConst{=}
       (\HOLTokenLambda{}\HOLBoundVar{E}.
            \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
            \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{CCS_Subst} \HOLBoundVar{E} \HOLBoundVar{t} \HOLBoundVar{X})) \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionscontextXXdef}{\UseVerbatim{HOLMultivariateDefinitionscontextXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsfromListXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps}. (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLSymConst{=} \HOLConst{FEMPTY} \HOLSymConst{|++} \HOLConst{ZIP} (\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsfromListXXdef}{\UseVerbatim{HOLMultivariateDefinitionsfromListXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsweaklyXXguardedXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLSymConst{=}
       (\HOLTokenLambda{}\HOLBoundVar{E}.
            \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
            \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{CCS_Subst} \HOLBoundVar{E} \HOLBoundVar{t} \HOLBoundVar{X})) \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsweaklyXXguardedXXdef}{\UseVerbatim{HOLMultivariateDefinitionsweaklyXXguardedXXdef}}
\newcommand{\HOLMultivariateDefinitions}{
\HOLDfnTag{Multivariate}{CCS_equation_def}\HOLMultivariateDefinitionsCCSXXequationXXdef
\HOLDfnTag{Multivariate}{CCS_solution_def}\HOLMultivariateDefinitionsCCSXXsolutionXXdef
\HOLDfnTag{Multivariate}{CCS_SUBST_def}\HOLMultivariateDefinitionsCCSXXSUBSTXXdef
\HOLDfnTag{Multivariate}{context_def}\HOLMultivariateDefinitionscontextXXdef
\HOLDfnTag{Multivariate}{fromList_def}\HOLMultivariateDefinitionsfromListXXdef
\HOLDfnTag{Multivariate}{weakly_guarded_def}\HOLMultivariateDefinitionsweaklyXXguardedXXdef
}
\begin{SaveVerbatim}{HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{BV} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E}) \HOLConst{\HOLTokenSubset{}}
       \HOLConst{BV} \HOLBoundVar{E} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} (\HOLConst{IMAGE} \HOLConst{BV} (\HOLConst{set} \HOLBoundVar{Ps}))
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}{\UseVerbatim{HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXsolutionXXlength}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{R} \HOLBoundVar{Ps}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{R} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXsolutionXXlength}{\UseVerbatim{HOLMultivariateTheoremsCCSXXsolutionXXlength}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXelim}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLSymConst{=} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXelim}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXelim}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{fm}. \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{FDOM} \HOLBoundVar{fm}) \HOLSymConst{\HOLTokenImp{}} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E} \HOLSymConst{=} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLConst{FEMPTY} \HOLBoundVar{E} \HOLSymConst{=} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} (\HOLTokenLambda{}(\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{p}). \HOLConst{FV} \HOLBoundVar{p} \HOLConst{\HOLTokenSubset{}} \HOLTokenLeftbrace{}\HOLBoundVar{x}\HOLTokenRightbrace{}) (\HOLConst{ZIP} (\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps})) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLSymConst{=}
        \HOLConst{FOLDR} (\HOLTokenLambda{}(\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{y}) \HOLBoundVar{e}. \HOLConst{CCS_Subst} \HOLBoundVar{e} \HOLBoundVar{y} \HOLBoundVar{x}) \HOLBoundVar{E} (\HOLConst{ZIP} (\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps})))
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{IS_PROC} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXnested}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       (\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{C}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Es}) \HOLBoundVar{C}) \HOLSymConst{=}
        \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLConst{MAP} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps})) \HOLBoundVar{Es}) \HOLBoundVar{C})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXnested}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXnested}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLConst{nil} \HOLSymConst{=} \HOLConst{nil}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXnil}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{=}
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXpar}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{=} \HOLBoundVar{u}\HOLSymConst{..}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXprefix}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXPROC}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{P}.
       (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{P}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{IS_PROC} \HOLBoundVar{P} \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{P} \HOLSymConst{=} \HOLBoundVar{P})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXPROC}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXPROC}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X} \HOLBoundVar{E}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} \HOLBoundVar{E}) \HOLSymConst{=}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} (\HOLConst{CCS_SUBST} (\HOLBoundVar{fm} \HOLSymConst{\HOLTokenBackslash{}\HOLTokenBackslash{}} \HOLBoundVar{X}) \HOLBoundVar{E})
       \HOLKeyword{else} \HOLConst{\ensuremath{\mu}} \HOLBoundVar{X} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrec}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXreduce}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{P} \HOLBoundVar{Ps}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLBoundVar{X} \HOLConst{\HOLTokenNotIn{}} \HOLConst{FV} \HOLBoundVar{e}) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}.
           \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} (\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs})) \HOLSymConst{\HOLTokenConj{}}
           (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLSymConst{=} \HOLBoundVar{E\sp{\prime}}) \HOLSymConst{\HOLTokenImp{}}
           (\HOLConst{CCS_SUBST} (\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLSymConst{=} \HOLConst{CCS_Subst} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{P} \HOLBoundVar{X})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXreduce}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXreduce}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{=} \HOLConst{relab} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}) \HOLBoundVar{rf}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrelab}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{=} \HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} (\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrestr}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXself}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{Xs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLConst{MAP} \HOLConst{var} \HOLBoundVar{Xs}) \HOLBoundVar{E} \HOLSymConst{=} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXself}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXself}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXSING}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}. \HOLConst{CCS_SUBST} (\HOLConst{FEMPTY} \HOLSymConst{|+} (\HOLBoundVar{X}\HOLSymConst{,}\HOLBoundVar{E\sp{\prime}})) \HOLBoundVar{E} \HOLSymConst{=} \HOLConst{CCS_Subst} \HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{X}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXSING}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXSING}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{=}
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXsum}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} (\HOLConst{var} \HOLBoundVar{X}) \HOLSymConst{=}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLBoundVar{fm} \HOLConst{'} \HOLBoundVar{X} \HOLKeyword{else} \HOLConst{var} \HOLBoundVar{X}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXvar}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXbackwardXXrules}
\HOLTokenTurnstile{} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXbackwardXXrules}{\UseVerbatim{HOLMultivariateTheoremscontextXXbackwardXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXcombin}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} (\HOLConst{context} \HOLBoundVar{Xs}) \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       (\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Es}) \HOLBoundVar{C})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXcombin}{\UseVerbatim{HOLMultivariateTheoremscontextXXcombin}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{context} \HOLBoundVar{Xs} \HOLConst{nil}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXnil}{\UseVerbatim{HOLMultivariateTheoremscontextXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXpar}{\UseVerbatim{HOLMultivariateTheoremscontextXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXparXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXparXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXparXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXparXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXparXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXparXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefix}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefixXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefixXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefixXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefixXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefixXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefixXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrec}{\UseVerbatim{HOLMultivariateTheoremscontextXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelab}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelabXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelabXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelabXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelabXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelabXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelabXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestr}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestrXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestrXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestrXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestrXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestrXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestrXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrules}
\HOLTokenTurnstile{} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{context} \HOLBoundVar{Xs} \HOLConst{nil}) \HOLSymConst{\HOLTokenConj{}} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E})) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrules}{\UseVerbatim{HOLMultivariateTheoremscontextXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsum}{\UseVerbatim{HOLMultivariateTheoremscontextXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsumXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsumXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXsumXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsumXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsumXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXsumXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{context} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXvar}{\UseVerbatim{HOLMultivariateTheoremscontextXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{e}) (\HOLConst{set} \HOLBoundVar{Xs})) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E})) (\HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}{\UseVerbatim{HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsdisjointXXimpXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}.
       \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsdisjointXXimpXXcontext}{\UseVerbatim{HOLMultivariateTheoremsdisjointXXimpXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}.
       \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}{\UseVerbatim{HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsEVERYXXweaklyXXguarded}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{X}. \HOLConst{MEM} \HOLBoundVar{E} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{CCS_Subst} \HOLBoundVar{E} \HOLBoundVar{t} \HOLBoundVar{X})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsEVERYXXweaklyXXguarded}{\UseVerbatim{HOLMultivariateTheoremsEVERYXXweaklyXXguarded}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFDOMXXfromList}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps}. (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}} (\HOLConst{FDOM} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLSymConst{=} \HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFDOMXXfromList}{\UseVerbatim{HOLMultivariateTheoremsFDOMXXfromList}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXEMPTY}
\HOLTokenTurnstile{} ([] \HOLSymConst{\HOLTokenMapto{}} []) \HOLSymConst{=} \HOLConst{FEMPTY}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXEMPTY}{\UseVerbatim{HOLMultivariateTheoremsfromListXXEMPTY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXEL}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{n} \HOLSymConst{\HOLTokenLt{}} \HOLConst{LENGTH} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       ((\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLConst{'} (\HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Xs}) \HOLSymConst{=} \HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Ps})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXEL}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXEL}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{P} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{n} \HOLSymConst{\HOLTokenLt{}} \HOLConst{LENGTH} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       ((\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}) \HOLConst{'} (\HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Xs}) \HOLSymConst{=} \HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Ps})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXHD}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{P} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       ((\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}) \HOLConst{'} \HOLBoundVar{X} \HOLSymConst{=} \HOLBoundVar{P})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXHD}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXHD}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXHD}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{xs} \HOLBoundVar{y} \HOLBoundVar{ys}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{x} \HOLBoundVar{xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{xs} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{ys}) \HOLSymConst{\HOLTokenImp{}}
       ((\HOLBoundVar{x}\HOLSymConst{::}\HOLBoundVar{xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{y}\HOLSymConst{::}\HOLBoundVar{ys}) \HOLSymConst{=} (\HOLBoundVar{xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{ys}) \HOLSymConst{|+} (\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{y}))
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXHD}{\UseVerbatim{HOLMultivariateTheoremsfromListXXHD}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{FV} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E}) \HOLConst{\HOLTokenSubset{}}
       \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} (\HOLConst{IMAGE} \HOLConst{FV} (\HOLConst{set} \HOLBoundVar{Ps}))
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}{\UseVerbatim{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{FV} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E}) \HOLConst{\HOLTokenSubset{}}
       \HOLConst{FV} \HOLBoundVar{E} \HOLConst{DIFF} \HOLConst{set} \HOLBoundVar{Xs} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} (\HOLConst{IMAGE} \HOLConst{FV} (\HOLConst{set} \HOLBoundVar{Ps}))
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}{\UseVerbatim{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsLISTXXRELXXequivalence}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}. \HOLConst{equivalence} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLConst{equivalence} (\HOLConst{LIST_REL} \HOLBoundVar{R})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsLISTXXRELXXequivalence}{\UseVerbatim{HOLMultivariateTheoremsLISTXXRELXXequivalence}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{OBS_CONGR} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}.
           \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{OBS_CONGR} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E})
             (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXreflexive}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}. \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Ps}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXreflexive}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXreflexive}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}.
           \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{OBS_contracts} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E})
             (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXtransitive}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLBoundVar{Rs}.
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{OBS_contracts} \HOLBoundVar{Qs} \HOLBoundVar{Rs} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Rs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXtransitive}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXtransitive}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{STRONG_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}.
           \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{STRONG_EQUIV} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E})
             (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsstrongXXuniqueXXsolution}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{Qs}.
           \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{STRONG_EQUIV} \HOLSymConst{\HOLTokenConj{}}
           \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{STRONG_EQUIV} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{STRONG_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsstrongXXuniqueXXsolution}{\UseVerbatim{HOLMultivariateTheoremsstrongXXuniqueXXsolution}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}.
           (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u} \HOLBoundVar{P\sp{\prime}}.
               \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLBoundVar{E\sp{\prime}} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}}
                   (\HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E\sp{\prime}}) \HOLSymConst{\HOLTokenConj{}}
                   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                       (\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
                       \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E}
                       \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd
                       \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}{\UseVerbatim{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{Qs}.
           \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLSymConst{\HOLTokenConj{}}
           \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{WEAK_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{Qs}.
           \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLSymConst{\HOLTokenConj{}}
           \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLSymConst{\HOLTokenImp{}}
           \HOLConst{OBS_CONGR} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{C}.
           \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenImp{}}
           (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l} \HOLBoundVar{R}.
                \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C} \HOLTokenWeakTransBegin\HOLConst{label} \HOLBoundVar{l}\HOLTokenImp{} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}}
                \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{C\sp{\prime}}.
                    \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                    \HOLBoundVar{R} \HOLConst{contracts} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                    (\HOLConst{WEAK_EQUIV} \HOLConst{O} (\HOLTokenLambda{}\HOLBoundVar{x} \HOLBoundVar{y}. \HOLBoundVar{x} \HOLTokenWeakTransBegin\HOLConst{label} \HOLBoundVar{l}\HOLTokenImp{} \HOLBoundVar{y}))
                      (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C})
                      (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C\sp{\prime}})) \HOLSymConst{\HOLTokenConj{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}.
               \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C} \HOLTokenWeakTransBegin\HOLConst{\ensuremath{\tau}}\HOLTokenImp{} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{C\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                   \HOLBoundVar{R} \HOLConst{contracts} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                   (\HOLConst{WEAK_EQUIV} \HOLConst{O} \HOLConst{EPS}) (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C})
                     (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C\sp{\prime}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{E} \HOLBoundVar{C\sb{\mathrm{0}}} \HOLBoundVar{Ps}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} (\HOLConst{context} \HOLBoundVar{Xs}) \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLFreeVar{C} \HOLSymConst{\HOLTokenConj{}}
       (\HOLBoundVar{E} \HOLSymConst{=} (\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ys})) \HOLBoundVar{Es})) \HOLSymConst{\HOLTokenConj{}}
       (\HOLBoundVar{C\sb{\mathrm{0}}} \HOLSymConst{=} (\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ys}) \HOLFreeVar{C})) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{n}. \HOLConst{OBS_contracts} (\HOLBoundVar{C\sb{\mathrm{0}}} \HOLBoundVar{Ps}) ((\HOLBoundVar{C\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}) \HOLBoundVar{Ps})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}.
           \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{u} \HOLBoundVar{P\sp{\prime}}.
               (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
               \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{e}) (\HOLConst{set} \HOLBoundVar{Xs})) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
               \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{C\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                   (\HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C\sp{\prime}}) \HOLSymConst{\HOLTokenConj{}}
                   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                       (\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
                       \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{E}
                       \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd
                       \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}}
       (\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       (\HOLBoundVar{E} \HOLSymConst{=} (\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ys})) \HOLBoundVar{Es})) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{x} \HOLBoundVar{P\sp{\prime}}.
           (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{e}) (\HOLConst{set} \HOLBoundVar{Xs})) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{E} \HOLBoundVar{Ps}) \HOLBoundVar{C} \HOLTokenTransBegin\HOLBoundVar{x}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{C\sp{\prime}}.
               \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C\sp{\prime}}) \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                   (\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
                   \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{E} \HOLBoundVar{Qs}) \HOLBoundVar{C}
                   \HOLTokenTransBegin\HOLBoundVar{x}\HOLTokenTransEnd
                   \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C} \HOLBoundVar{E} \HOLBoundVar{C\sb{\mathrm{0}}}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}}
       (\HOLBoundVar{E} \HOLSymConst{=} (\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ys})) \HOLBoundVar{Es})) \HOLSymConst{\HOLTokenConj{}}
       (\HOLBoundVar{C\sb{\mathrm{0}}} \HOLSymConst{=} (\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ys}) \HOLBoundVar{C})) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{n} \HOLBoundVar{xs} \HOLBoundVar{Ps} \HOLBoundVar{P\sp{\prime}}.
           (\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{EVERY} (\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{e}) (\HOLConst{set} \HOLBoundVar{Xs})) \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{TRACE} ((\HOLBoundVar{C\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}) \HOLBoundVar{Ps}) \HOLBoundVar{xs} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLBoundVar{xs} \HOLSymConst{\HOLTokenLeq{}} \HOLBoundVar{n} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{C\sp{\prime\prime}}.
               \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C\sp{\prime\prime}} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Ps}) \HOLBoundVar{C\sp{\prime\prime}}) \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                   (\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
                   \HOLConst{TRACE} ((\HOLBoundVar{C\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}) \HOLBoundVar{Qs}) \HOLBoundVar{xs}
                     (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Qs}) \HOLBoundVar{C\sp{\prime\prime}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}
\HOLTokenTurnstile{} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y}) \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXcombin}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       (\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{=} \HOLConst{LENGTH} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{CCS_SUBST} (\HOLBoundVar{Xs} \HOLSymConst{\HOLTokenMapto{}} \HOLBoundVar{Es}) \HOLBoundVar{C})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXcombin}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXcombin}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLConst{nil}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXnil}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXpar}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXprefix}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrec}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrecYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrecYY}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrecYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrecXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrecXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrecXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelab}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestr}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrules}
\HOLTokenTurnstile{} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLConst{nil}) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{u}\HOLSymConst{..}\HOLBoundVar{E})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{E\sb{\mathrm{2}}})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\nu}} \HOLBoundVar{L} \HOLBoundVar{E})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf})) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} (\HOLConst{FV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} (\HOLConst{BV} \HOLBoundVar{E}) (\HOLConst{set} \HOLBoundVar{Xs}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{\ensuremath{\mu}} \HOLBoundVar{Y} \HOLBoundVar{E})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrules}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsum}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLBoundVar{E\sb{\mathrm{2}}})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y}) \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXvar}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} (\HOLConst{var} \HOLBoundVar{Y})
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}}
\newcommand{\HOLMultivariateTheorems}{
\HOLThmTag{Multivariate}{BV_SUBSET_BIGUNION}\HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION
\HOLThmTag{Multivariate}{CCS_solution_length}\HOLMultivariateTheoremsCCSXXsolutionXXlength
\HOLThmTag{Multivariate}{CCS_SUBST_elim}\HOLMultivariateTheoremsCCSXXSUBSTXXelim
\HOLThmTag{Multivariate}{CCS_SUBST_elim'}\HOLMultivariateTheoremsCCSXXSUBSTXXelimYY
\HOLThmTag{Multivariate}{CCS_SUBST_FEMPTY}\HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY
\HOLThmTag{Multivariate}{CCS_SUBST_FOLDR}\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR
\HOLThmTag{Multivariate}{CCS_SUBST_IS_PROC}\HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC
\HOLThmTag{Multivariate}{CCS_SUBST_nested}\HOLMultivariateTheoremsCCSXXSUBSTXXnested
\HOLThmTag{Multivariate}{CCS_SUBST_nil}\HOLMultivariateTheoremsCCSXXSUBSTXXnil
\HOLThmTag{Multivariate}{CCS_SUBST_par}\HOLMultivariateTheoremsCCSXXSUBSTXXpar
\HOLThmTag{Multivariate}{CCS_SUBST_prefix}\HOLMultivariateTheoremsCCSXXSUBSTXXprefix
\HOLThmTag{Multivariate}{CCS_SUBST_PROC}\HOLMultivariateTheoremsCCSXXSUBSTXXPROC
\HOLThmTag{Multivariate}{CCS_SUBST_rec}\HOLMultivariateTheoremsCCSXXSUBSTXXrec
\HOLThmTag{Multivariate}{CCS_SUBST_reduce}\HOLMultivariateTheoremsCCSXXSUBSTXXreduce
\HOLThmTag{Multivariate}{CCS_SUBST_relab}\HOLMultivariateTheoremsCCSXXSUBSTXXrelab
\HOLThmTag{Multivariate}{CCS_SUBST_restr}\HOLMultivariateTheoremsCCSXXSUBSTXXrestr
\HOLThmTag{Multivariate}{CCS_SUBST_self}\HOLMultivariateTheoremsCCSXXSUBSTXXself
\HOLThmTag{Multivariate}{CCS_SUBST_SING}\HOLMultivariateTheoremsCCSXXSUBSTXXSING
\HOLThmTag{Multivariate}{CCS_SUBST_sum}\HOLMultivariateTheoremsCCSXXSUBSTXXsum
\HOLThmTag{Multivariate}{CCS_SUBST_var}\HOLMultivariateTheoremsCCSXXSUBSTXXvar
\HOLThmTag{Multivariate}{context_backward_rules}\HOLMultivariateTheoremscontextXXbackwardXXrules
\HOLThmTag{Multivariate}{context_combin}\HOLMultivariateTheoremscontextXXcombin
\HOLThmTag{Multivariate}{context_nil}\HOLMultivariateTheoremscontextXXnil
\HOLThmTag{Multivariate}{context_par}\HOLMultivariateTheoremscontextXXpar
\HOLThmTag{Multivariate}{context_par_rewrite}\HOLMultivariateTheoremscontextXXparXXrewrite
\HOLThmTag{Multivariate}{context_par_rule}\HOLMultivariateTheoremscontextXXparXXrule
\HOLThmTag{Multivariate}{context_prefix}\HOLMultivariateTheoremscontextXXprefix
\HOLThmTag{Multivariate}{context_prefix_rewrite}\HOLMultivariateTheoremscontextXXprefixXXrewrite
\HOLThmTag{Multivariate}{context_prefix_rule}\HOLMultivariateTheoremscontextXXprefixXXrule
\HOLThmTag{Multivariate}{context_rec}\HOLMultivariateTheoremscontextXXrec
\HOLThmTag{Multivariate}{context_relab}\HOLMultivariateTheoremscontextXXrelab
\HOLThmTag{Multivariate}{context_relab_rewrite}\HOLMultivariateTheoremscontextXXrelabXXrewrite
\HOLThmTag{Multivariate}{context_relab_rule}\HOLMultivariateTheoremscontextXXrelabXXrule
\HOLThmTag{Multivariate}{context_restr}\HOLMultivariateTheoremscontextXXrestr
\HOLThmTag{Multivariate}{context_restr_rewrite}\HOLMultivariateTheoremscontextXXrestrXXrewrite
\HOLThmTag{Multivariate}{context_restr_rule}\HOLMultivariateTheoremscontextXXrestrXXrule
\HOLThmTag{Multivariate}{context_rules}\HOLMultivariateTheoremscontextXXrules
\HOLThmTag{Multivariate}{context_sum}\HOLMultivariateTheoremscontextXXsum
\HOLThmTag{Multivariate}{context_sum_rewrite}\HOLMultivariateTheoremscontextXXsumXXrewrite
\HOLThmTag{Multivariate}{context_sum_rule}\HOLMultivariateTheoremscontextXXsumXXrule
\HOLThmTag{Multivariate}{context_var}\HOLMultivariateTheoremscontextXXvar
\HOLThmTag{Multivariate}{DISJOINT_BV_CCS_SUBST}\HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST
\HOLThmTag{Multivariate}{disjoint_imp_context}\HOLMultivariateTheoremsdisjointXXimpXXcontext
\HOLThmTag{Multivariate}{disjoint_imp_weakly_guarded}\HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded
\HOLThmTag{Multivariate}{EVERY_weakly_guarded}\HOLMultivariateTheoremsEVERYXXweaklyXXguarded
\HOLThmTag{Multivariate}{FDOM_fromList}\HOLMultivariateTheoremsFDOMXXfromList
\HOLThmTag{Multivariate}{fromList_EMPTY}\HOLMultivariateTheoremsfromListXXEMPTY
\HOLThmTag{Multivariate}{fromList_FAPPLY_EL}\HOLMultivariateTheoremsfromListXXFAPPLYXXEL
\HOLThmTag{Multivariate}{fromList_FAPPLY_EL'}\HOLMultivariateTheoremsfromListXXFAPPLYXXELYY
\HOLThmTag{Multivariate}{fromList_FAPPLY_HD}\HOLMultivariateTheoremsfromListXXFAPPLYXXHD
\HOLThmTag{Multivariate}{fromList_HD}\HOLMultivariateTheoremsfromListXXHD
\HOLThmTag{Multivariate}{FV_SUBSET_BIGUNION}\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION
\HOLThmTag{Multivariate}{FV_SUBSET_BIGUNION_PRO}\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO
\HOLThmTag{Multivariate}{LIST_REL_equivalence}\HOLMultivariateTheoremsLISTXXRELXXequivalence
\HOLThmTag{Multivariate}{OBS_CONGR_subst_context}\HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext
\HOLThmTag{Multivariate}{OBS_contracts_reflexive}\HOLMultivariateTheoremsOBSXXcontractsXXreflexive
\HOLThmTag{Multivariate}{OBS_contracts_subst_context}\HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext
\HOLThmTag{Multivariate}{OBS_contracts_transitive}\HOLMultivariateTheoremsOBSXXcontractsXXtransitive
\HOLThmTag{Multivariate}{STRONG_EQUIV_subst_context}\HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext
\HOLThmTag{Multivariate}{strong_unique_solution}\HOLMultivariateTheoremsstrongXXuniqueXXsolution
\HOLThmTag{Multivariate}{strong_unique_solution_lemma}\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma
\HOLThmTag{Multivariate}{unique_solution_of_obs_contractions}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions
\HOLThmTag{Multivariate}{unique_solution_of_rooted_contractions}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions
\HOLThmTag{Multivariate}{unique_solution_of_rooted_contractions_lemma}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma
\HOLThmTag{Multivariate}{USC_unfolding_lemma1}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne
\HOLThmTag{Multivariate}{USC_unfolding_lemma2}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo
\HOLThmTag{Multivariate}{USC_unfolding_lemma3}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree
\HOLThmTag{Multivariate}{USC_unfolding_lemma4}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour
\HOLThmTag{Multivariate}{weakly_guarded_backward_rules}\HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules
\HOLThmTag{Multivariate}{weakly_guarded_combin}\HOLMultivariateTheoremsweaklyXXguardedXXcombin
\HOLThmTag{Multivariate}{weakly_guarded_imp_context}\HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext
\HOLThmTag{Multivariate}{weakly_guarded_nil}\HOLMultivariateTheoremsweaklyXXguardedXXnil
\HOLThmTag{Multivariate}{weakly_guarded_par}\HOLMultivariateTheoremsweaklyXXguardedXXpar
\HOLThmTag{Multivariate}{weakly_guarded_par_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_par_rule}\HOLMultivariateTheoremsweaklyXXguardedXXparXXrule
\HOLThmTag{Multivariate}{weakly_guarded_prefix}\HOLMultivariateTheoremsweaklyXXguardedXXprefix
\HOLThmTag{Multivariate}{weakly_guarded_prefix_rule}\HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule
\HOLThmTag{Multivariate}{weakly_guarded_rec}\HOLMultivariateTheoremsweaklyXXguardedXXrec
\HOLThmTag{Multivariate}{weakly_guarded_rec'}\HOLMultivariateTheoremsweaklyXXguardedXXrecYY
\HOLThmTag{Multivariate}{weakly_guarded_rec_rule}\HOLMultivariateTheoremsweaklyXXguardedXXrecXXrule
\HOLThmTag{Multivariate}{weakly_guarded_relab}\HOLMultivariateTheoremsweaklyXXguardedXXrelab
\HOLThmTag{Multivariate}{weakly_guarded_relab_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_relab_rule}\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule
\HOLThmTag{Multivariate}{weakly_guarded_restr}\HOLMultivariateTheoremsweaklyXXguardedXXrestr
\HOLThmTag{Multivariate}{weakly_guarded_restr_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_restr_rule}\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule
\HOLThmTag{Multivariate}{weakly_guarded_rules}\HOLMultivariateTheoremsweaklyXXguardedXXrules
\HOLThmTag{Multivariate}{weakly_guarded_sum}\HOLMultivariateTheoremsweaklyXXguardedXXsum
\HOLThmTag{Multivariate}{weakly_guarded_sum_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_sum_rule}\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule
\HOLThmTag{Multivariate}{weakly_guarded_var}\HOLMultivariateTheoremsweaklyXXguardedXXvar
\HOLThmTag{Multivariate}{weakly_guarded_var_rule}\HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule
}