### 2019-09-09 ic-journal/equa.tex

\section{Equations and contractions}
\label{s:eq}

\hl{In the CCS syntax given in previous section, there is the syntactic
possibility that an agent variable $A$ occurs
outside the recursion operator of the same variable (or, equivalently, a
process constant $C$ is used but undefined). We say such a variable is
\emph{free}, while those variables wrapped by their recursions are
\emph{bound}.}
\hl{A CCS \emph{tem} (or \emph{expression}, i.e. those with valid syntax)
without any free variable is called a CCS \emph{process}.}
\hl{For instance, the variable $X$ is free in $a.X + \recu Y (b.Y)$, while $Y$
is bound. There is also the syntactic possibility that the same
variable $X$ is
both free and bound, e.g. in $a.X + \recu X (b.X)$.
This does not cause any ambiguity, but sometimes leads to more
complex proofs.} (c.f. Section~\ref{sec:multivariate} for more details.)

\hl{In this paper (and the formalisation work), we reuse the free agent
variables as \emph{equation variables}. This eliminates the need of
another type for CCS equations, and we can also reuse the existing
variable substitution operation} (c.f. SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) \hl{for the substitution of equation variables.
For example, the result of substituting equation variable $X$ to $\nil$ in $a.X +
\recu X (b.X)$, denoted by $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ with the part $\recu X (b.X)$
untouched. Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$. (and $E[P]$ for $E \sub P X$ if there is a single equation variable $X$.)}

\hl{In fact, free agent variables have the same transitional behavior
as the deadlock $\nil$ according to the SOS rules, as there is
no rule for their transitions at all. In fact, most CCS theorems still hold if
the involved CCS terms contain free variables. (The most notable
exceptions are all versions of unique solution of equations/contracts,
where all solutions must be \emph{pure} processes (i.e. no free variable).)}

\subsection{Systems of equations}
\label{ss:SysEq}

Milner's ``unique solution of equations'' theorems~\cite{Mil89} intuitively
say that, if a context $C$ \hl{(i.e., a CCS expression with possibly
free variables)}\footnote{\hl{Rigorously speaking, under our setting
(i.e.~reusing free agent variables as equation variables) not all
valid CCS expressions
are valid contexts: those where free variables occur inside
recusion operators must be all excluded. For instace, $\recu X (a.X +
b.Y)$ is not a valid context with the variable $Y$. This special requirement
is perfectly aligned with CCS literature using process constants, as any
equation variable cannot occur inside the definition of any
constant. In fact, all versions of ``unique solution of
equations/contractions'' theorems do not hold if equation variables
are allowed to occur inside any recursion operator.}} obeys certain conditions,
then all processes $P$ that satisfy the equation $P \wb \ct P$ are
bisimilar with each other.

\begin{definition}[equations] % Def 3.1
  \label{def:equation}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ containing variables up to $\til X$. Then 
$\{ X_i = E_i\}_{i\in I}$ is 
  a \emph{system of equations}. (There is one equation $E_i$ for each variable $X_i$.)
\end{definition}

The components of $\til P$ need not be
different from each other, as it must be for the variables $\til X$.
If the system has infinitely many equations, the  tuples $\til P$ and $\til X$ are infinite too.

\begin{definition}[solutions and unique solutions]
  \label{def:solution}
Suppose $\{ X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the system of equations (for $\wb$)} 
if for each $i$ it holds that $P_i \wb E_i [\til P]$;
\item The system has \emph{a unique solution for $\wb$}  if whenever 
 $\til P$ and $\til Q$ are both solutions \hl{it holds that} $\til P \wb \til Q$. 
\end{itemize} 
 \end{definition}
\hl{Similarily, the \emph{(unique) solution of a system of equations for $\sim$}
(or for $\rapprox$) can be obtained by replacing all occurrences of $\wb$
in above definition with $\sim$ and $rapprox$, respectively.}

For instance, the solution of the equation $X = a. X$ 
is the process
$R \DSdefi \recu A {\, (a. A)}$, and for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, for instance,
 $K$ and $K | b$, for $K \DSdefi \recu K {\, (a. K)}$. (Actually any process capable of
continuously performing \hl{$a$--actions} is a solution \hl{of} $X = a|  X$.)

%
% The unique solution of the system (1), modulo $\wb$,  is the constant $K \Defi a
% . K$:  for any other solution $P$ we have $P \wb K$.
% The unique solution of (2), modulo $\wb$, are the constants $K_1 , K_2$
% with $K_1 \Defi a . K_2$ and $K_2 \Defi b. K_1$; again, for any other
% pair of solutions $P_1,P_2$ we have $K_1 \wb P_1$ and $K_2 \wb P_2$.
%
Examples of systems that do not have unique solutions are: $X = X$, $X
= \tau . X$ and $X = a | X$.

\begin{definition}[guardedness of equations]
\label{def:guardness}
A system of equations $\{ X_i = E_i\}_{i\in I}$ is 
\begin{itemize}
\item \emph{weakly guarded} if, in each $E_i$, each occurrence of
  each $X_i$ is underneath a prefix;

\item \emph{(strongly) guarded} if, in each $E_i$, each occurrence of
  each $X_i$ is underneath a \emph{visible} prefix;

\item \emph{sequential} if, in each $E_i$, each
  \hl{occurrence of each $X_i$ is only underneath prefixes and sums.}
\end{itemize}
\end{definition}

\hl{In other words, if a system of equations is sequential, then for
each context $E_i$, any subcontext of $E_i$ in which $X_j $
appears, apart from $X_j$ itself, is a sum of prefixed expressions.
For instance,}
\begin{itemize}
\item $X = \tau. X + \mu . \nil$ is sequential but not 
 guarded, because the guarding prefix for the variable
is not visible;
\item $X =  \ell . X | P$ is guarded but not sequential;
\item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as
$X = \tau . (a. X + \tau . b .X + \tau  )$ are both guarded and sequential.
\end{itemize}

\hl{Milner has totally three versions of ``unique solution of equations''
theorems, for $\sim$, $\wb$ and $\rapprox$, respectively, but only the
following two versions are explicitly mentioned} in~\citep[p.~103, 158]{Mil89}:
\begin{theorem}[unique solution of equations for $\sim$]
\label{t:Mil89s1}
Let $E_i$ be weakly guarded with free variables up to $\til X$,
and let ${\til P} \sim {\til E}\{\til P /\til X\}$,
  ${\til Q} \sim {\til E}\{\til Q /\til X\}$. Then ${\til P} \sim {\til Q}$.
\end{theorem}

\begin{theorem}[unique solution of equations for $\rapprox$]
\label{t:Mil89s3}
Let $E_i$ be guarded and sequential with free
variables up to $\til X$, and let ${\til P} \rapprox {\til E}\{\til P /\til X\}$,
  ${\til Q} \rapprox {\til E}\{\til Q /\til X\}$. Then ${\til P} \rapprox {\til Q}$.
\end{theorem}

\hl{The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums where equation variables appear are guarded
sums}\footnote{But if the CCS syntax were defined with only guarded
  sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
  in~\cite{sangiorgi2015equations}, this addition
  requirement disappears automatically, and we can even say $\wb$ is indeed a
  congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded (with guarded sums only) and sequential with free
variables up to $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

The proof of \hl{the last theorem above} exploits an invariance
property on immediate \hl{derivations}
of guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

\hl{For more details of above three theorems,} c.f. Section~\ref{ss:part2}
for their formalised versions (the univariate case) and
Section~\ref{sec:multivariate} for the multivariate case
(Theorem~\ref{t:Mil89s1} only).

%% next file: contraction.tex