### 2019-09-15 ic-journal/intro.tex

\section{Introduction}

A prominent proof method for bisimulation, put forward by Robin Milner and widely used in his
landmark CCS book~\cite{Mil89}, is the
\emph{unique solution of equations}, whereby two tuples of processes are
componentwise bisimilar if they are solutions of the same system of equations.
This method is important in verification techniques and tools
based on algebraic reasoning~\cite{BaeBOOK,theoryAndPractice,RosUnder10}.

\hl{Milner's unique-solution theorem for weak bisimilarity, however,
has severe syntactic limitations:}
the equations must be both \emph{guarded} and \emph{sequential}. That is,
variables of the equations \hl{can only occur underneath visible prefixes and summations.}
One way \hl{to overcome} such limitations is to replace the equations
with special inequations called
\emph{contractions}~\cite{sangiorgi2015equations,sangiorgi2017equations}. Contraction is a
preorder that, roughly, places some efficiency constraints on processes.
\hl{The unique solution of contractions is defined as with equations:}
any two solutions must be componentwise bisimilar.
The difference \hl{from} equations is \hl{in} the meaning of a solution:
in the case of contractions the solution is evaluated with respect to
the contraction preorder, rather than bisimilarity. 
With contractions, most syntactic limitations of the unique-solution theorem \hl{are
eliminated}. One constraint that still remains
in~\cite{sangiorgi2017equations}
(where the issue is bypassed using a more
restrictive CCS syntax)
is the occurrences of %\emph{direct} 
\hl{sums} (e.g.~$P + Q$) due to
the \hl{non-substitutivity of the contraction preorder in this case}.

 \hl{This paper presents}  a
comprehensive formalisation of Milner's Calculus of Communicating
Systems (CCS) in the HOL theorem prover (HOL4),
with a focus towards the theory of unique solutions of equations and contractions.
\hl{Many results in Milner's CCS book}~\cite{Mil89} are covered, since
the unique-solution theorems rely on a large number of fundamental results.
Indeed the formalisation encompasses all basic properties of strong and weak
\hl{bisimilarities} (e.g.~the fixed-point and substitutivity properties), the
basic properties of rooted bisimilarity and their algebraic laws.
Further extensions include several versions of ``bisimulation up to''
techniques, the expansion and contraction preorder.
Concerning rooted bisimilarity, the formalisation
includes Hennessy's Lemma and Deng's Lemma,
 and two theorems saying that rooted bisimilarity is the coarsest (largest)
 congruence contained in weak bisimilarity ($\wb$): one is classical
 with the hypothesis that no process uses up all labels;
the other without such hypothesis, essentially formalised van Glabbeek's proof~\cite{van2005characterisation}.
Similar results are also proved for the rooted contraction preorder.
In this respect, the work is also an extensive experiment using the HOL theorem prover with its
recent developments, including its coinduction defining package.

From the point of view of the CCS theory, this formalisation has offered us the possibility of
further refining the theory of unique solutions of contractions.
In particular, the existing result~\cite{sangiorgi2017equations} has limitations on the body of the contractions due to the
substitutivity problems of weak bisimilarity and other behavioural relations with respect
to the sum operator.
In this paper, the contraction-based proof technique is further
refined by moving from the contraction preorder to
\emph{rooted contraction}, that is, the coarsest precongruence contained in the contraction
preorder. The resulting unique-solution theorem is now valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself), and places no 
\hl{constraints on the use the sum operator}.

Another benefit of the formalisation is 
that we can take advantage of results about different 
equivalences and preorders that share similar proof structures.
Such structural simililarities can be found, for instance, in the
following cases: the proofs that rooted bisimilarity and rooted
contraction are, respectively, the coarsest congruence contained in
weak bisimilarity and the coarsest precongruence contained in the
contraction preorder; the proofs about unique solution(s) of equations
for weak bisimilarity that uses the contraction preorder as an
auxiliary relation, and other unique solution results (e.g. the one
for rooted in which the auxiliary relation is rooted contraction); the
proofs about various forms of enhancements of the bisimulation proof
method (the `up-to' techniques).
%
In these cases, when moving between proofs there are only a few places in
which the HOL proof scripts have to be modified.
Then the successful termination of the proof gives us a guarantee that the proof is
complete and trustworthy, eliminating the risks 
of overlooking or missing details as in \emph{paper-and-pencil} proofs.

We first consider the case of a single equation (or contraction),
that we call the \univariate case, then consider the \multivariate
  case, where  more than one equation
  (or contraction) \hl{with multiple equation variables is considered.}
The \univariate theorems use
$\lambda$-functions to present CCS equations,
while the \multivariate versions require a
more careful and delicate treatment of CCS
expressions with multiple variables and substitutions acting on them.

In contrast to modern literature, e.g.~\cite{Gorrieri:2015jt},
we have followed Milner's original approach~\cite{milner1990operational} and adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
\emph{process} is a CCS term without equation variables. 
This allows us a smoother move from the \univariate case to the \multivariate
one. See Section~\ref{sec:multivariate} for  details.

\paragraph{Structure of the paper}

Section~\ref{ss:ccs} presents basic background materials on CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Section~\ref{s:eq} discusses equations and contractions.
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
Section~\ref{s:for} highlights our formalisation in HOL4 for the
\univariate case.
Section~\ref{sec:multivariate} describes the \hl{extension} to the
  \multivariate case.
Finally,  Section~\ref{s:rel} and \ref{s:concl} discuss related work,
conclusions, and a few directions for future work.

%%% next file: ccs.tex