### 2019-11-26 paper/HOLMultivariate.tex
Re-generated theories (I take the lock back and sync all changes from .tex to .htex)

\newcommand{\HOLMultivariateDate}{19 Novembre 2019}
\newcommand{\HOLMultivariateTime}{10:23}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXequationXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Es}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXequationXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXequationXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXsolutionXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps}.
       \HOLConst{CCS_solution} \HOLBoundVar{R} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{LIST_REL} \HOLBoundVar{R} \HOLBoundVar{Ps} \ensuremath{(}\HOLConst{MAP} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{Es}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXsolutionXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXsolutionXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsCCSXXSUBSTXXdef}
\HOLTokenTurnstile{} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{nu} \HOLBoundVar{L} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E} \HOLBoundVar{rf}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{relab} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X}.
        \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{var} \HOLBoundVar{X}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
        \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLBoundVar{fm} \HOLConst{'} \HOLBoundVar{X} \HOLKeyword{else} \HOLConst{var} \HOLBoundVar{X}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X} \HOLBoundVar{E}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{X} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLBoundVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLBoundVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLBoundVar{X}\ensuremath{)} \HOLBoundVar{E}\ensuremath{)}
       \HOLKeyword{else} \HOLConst{rec} \HOLBoundVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsCCSXXSUBSTXXdef}{\UseVerbatim{HOLMultivariateDefinitionsCCSXXSUBSTXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionscontextXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Xs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionscontextXXdef}{\UseVerbatim{HOLMultivariateDefinitionscontextXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsfromListXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps}. \HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{FEMPTY} \HOLSymConst{|++} \HOLConst{ZIP} \ensuremath{(}\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsfromListXXdef}{\UseVerbatim{HOLMultivariateDefinitionsfromListXXdef}}
\begin{SaveVerbatim}{HOLMultivariateDefinitionsweaklyXXguardedXXdef}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Xs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateDefinitionsweaklyXXguardedXXdef}{\UseVerbatim{HOLMultivariateDefinitionsweaklyXXguardedXXdef}}
\newcommand{\HOLMultivariateDefinitions}{
\HOLDfnTag{Multivariate}{CCS_equation_def}\HOLMultivariateDefinitionsCCSXXequationXXdef
\HOLDfnTag{Multivariate}{CCS_solution_def}\HOLMultivariateDefinitionsCCSXXsolutionXXdef
\HOLDfnTag{Multivariate}{CCS_SUBST_def}\HOLMultivariateDefinitionsCCSXXSUBSTXXdef
\HOLDfnTag{Multivariate}{context_def}\HOLMultivariateDefinitionscontextXXdef
\HOLDfnTag{Multivariate}{fromList_def}\HOLMultivariateDefinitionsfromListXXdef
\HOLDfnTag{Multivariate}{weakly_guarded_def}\HOLMultivariateDefinitionsweaklyXXguardedXXdef
}
\begin{SaveVerbatim}{HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{BV} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}{\UseVerbatim{HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXsolutionXXlength}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{R} \HOLBoundVar{Ps}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CCS_solution} \HOLBoundVar{R} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXsolutionXXlength}{\UseVerbatim{HOLMultivariateTheoremsCCSXXsolutionXXlength}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXelim}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXelim}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXelim}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{fm}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{FDOM} \HOLBoundVar{fm}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXelimYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLConst{FEMPTY} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{p}\ensuremath{)}. \HOLConst{FV} \HOLBoundVar{p} \HOLConst{\HOLTokenSubset{}} \HOLTokenLeftbrace{}\HOLBoundVar{x}\HOLTokenRightbrace{}\ensuremath{)} \ensuremath{(}\HOLConst{ZIP} \ensuremath{(}\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{FOLDR} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{y}\ensuremath{)} \HOLBoundVar{e}. \ensuremath{[}\HOLBoundVar{y}\ensuremath{/}\HOLBoundVar{x}\ensuremath{]} \HOLBoundVar{e}\ensuremath{)} \HOLBoundVar{E} \ensuremath{(}\HOLConst{ZIP} \ensuremath{(}\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDRYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{FEVERY} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{p}\ensuremath{)}. \HOLConst{FV} \HOLBoundVar{p} \HOLConst{\HOLTokenSubset{}} \HOLTokenLeftbrace{}\HOLBoundVar{x}\HOLTokenRightbrace{}\ensuremath{)} \ensuremath{(}\HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{FOLDR} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{y}\ensuremath{)} \HOLBoundVar{e}. \ensuremath{[}\HOLBoundVar{y}\ensuremath{/}\HOLBoundVar{x}\ensuremath{]} \HOLBoundVar{e}\ensuremath{)} \HOLBoundVar{E} \ensuremath{(}\HOLConst{ZIP} \ensuremath{(}\HOLBoundVar{Xs}\HOLSymConst{,}\HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDRYY}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXFOLDRYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{IS_PROC} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXnested}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Es}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLConst{MAP} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{Es}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXnested}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXnested}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXnil}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXpar}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXprefix}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXPROC}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{P}.
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{P}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{IS_PROC} \HOLBoundVar{P} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{P} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{P}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXPROC}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXPROC}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X} \HOLBoundVar{E}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{X} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLBoundVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLBoundVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLBoundVar{X}\ensuremath{)} \HOLBoundVar{E}\ensuremath{)}
       \HOLKeyword{else} \HOLConst{rec} \HOLBoundVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrec}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXreduce}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{P} \HOLBoundVar{Ps}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLBoundVar{X} \HOLConst{\HOLTokenNotIn{}} \HOLConst{FV} \HOLBoundVar{e}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \ensuremath{[}\HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{P}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXreduce}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXreduce}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{relab} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)} \HOLBoundVar{rf}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrelab}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{nu} \HOLBoundVar{L} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLBoundVar{fm} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXrestr}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXself}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{Xs}. \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{(}\ensuremath{[}\HOLConst{MAP} \HOLConst{var} \HOLBoundVar{Xs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXself}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXself}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXsing}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \ensuremath{[}\ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{]}\ensuremath{/}\ensuremath{[}\HOLBoundVar{X}\ensuremath{]}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXsing}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXsing}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXSING}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{CCS_SUBST} \ensuremath{(}\HOLConst{FEMPTY} \HOLSymConst{|+} \ensuremath{(}\HOLBoundVar{X}\HOLSymConst{,}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXSING}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXSING}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXsum}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsCCSXXSUBSTXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{fm} \HOLBoundVar{X}.
       \HOLConst{CCS_SUBST} \HOLBoundVar{fm} \ensuremath{(}\HOLConst{var} \HOLBoundVar{X}\ensuremath{)} \HOLSymConst{\ensuremath{=}}
       \HOLKeyword{if} \HOLBoundVar{X} \HOLConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLBoundVar{fm} \HOLKeyword{then} \HOLBoundVar{fm} \HOLConst{'} \HOLBoundVar{X} \HOLKeyword{else} \HOLConst{var} \HOLBoundVar{X}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsCCSXXSUBSTXXvar}{\UseVerbatim{HOLMultivariateTheoremsCCSXXSUBSTXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXbackwardXXrules}
\HOLTokenTurnstile{} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{Y} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E} \HOLConst{DELETE} \HOLBoundVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXbackwardXXrules}{\UseVerbatim{HOLMultivariateTheoremscontextXXbackwardXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXcombin}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLConst{context} \HOLBoundVar{Xs}\ensuremath{)} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Es}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXcombin}{\UseVerbatim{HOLMultivariateTheoremscontextXXcombin}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{context} \HOLBoundVar{Xs} \HOLConst{\ensuremath{\mathbf{0}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXnil}{\UseVerbatim{HOLMultivariateTheoremscontextXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXpar}{\UseVerbatim{HOLMultivariateTheoremscontextXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXparXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXparXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXparXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXparXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXparXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXparXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefix}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefixXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefixXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefixXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXprefixXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXprefixXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXprefixXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{Y} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E} \HOLConst{DELETE} \HOLBoundVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrec}{\UseVerbatim{HOLMultivariateTheoremscontextXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelab}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelabXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelabXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelabXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrelabXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrelabXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXrelabXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestr}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestrXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestrXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestrXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrestrXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrestrXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXrestrXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXrules}
\HOLTokenTurnstile{} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{context} \HOLBoundVar{Xs} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXrules}{\UseVerbatim{HOLMultivariateTheoremscontextXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsum}{\UseVerbatim{HOLMultivariateTheoremscontextXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsumXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsumXXrewrite}{\UseVerbatim{HOLMultivariateTheoremscontextXXsumXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXsumXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXsumXXrule}{\UseVerbatim{HOLMultivariateTheoremscontextXXsumXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremscontextXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremscontextXXvar}{\UseVerbatim{HOLMultivariateTheoremscontextXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}{\UseVerbatim{HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsdisjointXXimpXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsdisjointXXimpXXcontext}{\UseVerbatim{HOLMultivariateTheoremsdisjointXXimpXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}{\UseVerbatim{HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsEVERYXXweaklyXXguarded}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{X}. \HOLConst{MEM} \HOLBoundVar{E} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsEVERYXXweaklyXXguarded}{\UseVerbatim{HOLMultivariateTheoremsEVERYXXweaklyXXguarded}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFDOMXXfromList}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps}.
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{FDOM} \ensuremath{(}\HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFDOMXXfromList}{\UseVerbatim{HOLMultivariateTheoremsFDOMXXfromList}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXDOMSUBXXNOTXXINXXDOM}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{Ps}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{\setminus}} \HOLBoundVar{X} \HOLSymConst{\ensuremath{=}} \HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXDOMSUBXXNOTXXINXXDOM}{\UseVerbatim{HOLMultivariateTheoremsfromListXXDOMSUBXXNOTXXINXXDOM}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXEMPTY}
\HOLTokenTurnstile{} \HOLConst{fromList} \ensuremath{[}\ensuremath{]} \ensuremath{[}\ensuremath{]} \HOLSymConst{\ensuremath{=}} \HOLConst{FEMPTY}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXEMPTY}{\UseVerbatim{HOLMultivariateTheoremsfromListXXEMPTY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXEL}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{n} \HOLSymConst{\HOLTokenLt{}} \HOLConst{LENGTH} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLConst{'} \ensuremath{(}\HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Ps}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXEL}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXEL}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{P} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{n} \HOLSymConst{\HOLTokenLt{}} \HOLConst{LENGTH} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{fromList} \ensuremath{(}\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs}\ensuremath{)} \ensuremath{(}\HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}\ensuremath{)} \HOLConst{'} \ensuremath{(}\HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{EL} \HOLBoundVar{n} \HOLBoundVar{Ps}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXELYY}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXFAPPLYXXHD}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{P} \HOLBoundVar{Ps} \HOLBoundVar{n}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{fromList} \ensuremath{(}\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs}\ensuremath{)} \ensuremath{(}\HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}\ensuremath{)} \HOLConst{'} \HOLBoundVar{X} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{P}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXFAPPLYXXHD}{\UseVerbatim{HOLMultivariateTheoremsfromListXXFAPPLYXXHD}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsfromListXXHD}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{Xs} \HOLBoundVar{P} \HOLBoundVar{Ps}.
       \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{X} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \ensuremath{(}\HOLConst{fromList} \ensuremath{(}\HOLBoundVar{X}\HOLSymConst{::}\HOLBoundVar{Xs}\ensuremath{)} \ensuremath{(}\HOLBoundVar{P}\HOLSymConst{::}\HOLBoundVar{Ps}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLConst{fromList} \HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLSymConst{|+} \ensuremath{(}\HOLBoundVar{X}\HOLSymConst{,}\HOLBoundVar{P}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsfromListXXHD}{\UseVerbatim{HOLMultivariateTheoremsfromListXXHD}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{FV} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{FV} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}{\UseVerbatim{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{FV} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLConst{\HOLTokenSubset{}}
       \HOLConst{FV} \HOLBoundVar{E} \HOLConst{DIFF} \HOLConst{set} \HOLBoundVar{Xs} \HOLConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{FV} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Ps}\ensuremath{)}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}{\UseVerbatim{HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsLISTXXRELXXequivalence}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}. \HOLConst{equivalence} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLConst{equivalence} \ensuremath{(}\HOLConst{LIST_REL} \HOLBoundVar{R}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsLISTXXRELXXequivalence}{\UseVerbatim{HOLMultivariateTheoremsLISTXXRELXXequivalence}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{OBS_CONGR} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{OBS_CONGR} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXreflexive}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}. \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Ps}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXreflexive}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXreflexive}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{OBS_contracts} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsOBSXXcontractsXXtransitive}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLBoundVar{Rs}.
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{OBS_contracts} \HOLBoundVar{Qs} \HOLBoundVar{Rs} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{OBS_contracts} \HOLBoundVar{Ps} \HOLBoundVar{Rs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsOBSXXcontractsXXtransitive}{\UseVerbatim{HOLMultivariateTheoremsOBSXXcontractsXXtransitive}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{STRONG_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{STRONG_EQUIV} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}{\UseVerbatim{HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLBoundVar{E} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}.
           \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
               \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{E}\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLConst{\HOLTokenSubset{}} \HOLConst{set} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}}
                   \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                   \ensuremath{(}\ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
                       \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{E}\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}{\UseVerbatim{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXthm}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{STRONG_EQUIV} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{STRONG_EQUIV} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{STRONG_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXthm}{\UseVerbatim{HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXthm}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{WEAK_EQUIV} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Ps} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLBoundVar{Qs} \HOLConst{\HOLTokenIn{}} \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{OBS_CONGR} \HOLBoundVar{Ps} \HOLBoundVar{Qs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLBoundVar{Qs}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Qs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{C}.
           \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
           \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l} \HOLBoundVar{R}.
                \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C} \HOLTokenWeakTransBegin\HOLConst{label} \HOLBoundVar{l}\HOLTokenWeakTransEnd \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}}
                \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{C}\sp{\prime}}.
                    \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                    \HOLBoundVar{R} \HOLConst{contracts} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                    \ensuremath{(}\HOLConst{WEAK_EQUIV} \HOLConst{O} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{x} \HOLBoundVar{y}. \HOLBoundVar{x} \HOLTokenWeakTransBegin\HOLConst{label} \HOLBoundVar{l}\HOLTokenWeakTransEnd \HOLBoundVar{y}\ensuremath{)}\ensuremath{)}
                      \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}.
               \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C} \HOLTokenWeakTransBegin\HOLConst{\ensuremath{\tau}}\HOLTokenWeakTransEnd \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{C}\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                   \HOLBoundVar{R} \HOLConst{contracts} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
                   \ensuremath{(}\HOLConst{WEAK_EQUIV} \HOLConst{O} \HOLConst{EPS}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}{\UseVerbatim{HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLBoundVar{Ps}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLConst{context} \HOLBoundVar{Xs}\ensuremath{)} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{CCS_solution} \HOLConst{OBS_contracts} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLFreeVar{C} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} \ensuremath{[}\HOLBoundVar{Ys}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{Es}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLSymConst{\ensuremath{=}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{Ys}. \ensuremath{[}\HOLBoundVar{Ys}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{n}. \HOLConst{OBS_contracts} \ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLBoundVar{Ps}\ensuremath{)} \ensuremath{(}\ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}\ensuremath{)} \HOLBoundVar{Ps}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}.
           \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
               \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
               \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
               \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
               \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{C}\sp{\prime}}.
                   \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                   \ensuremath{(}\ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
                   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
                       \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C} \HOLBoundVar{E}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} \ensuremath{[}\HOLBoundVar{Ys}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{Es}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps} \HOLBoundVar{x} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
           \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \ensuremath{[}\HOLBoundVar{E} \HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C} \HOLTokenTransBegin\HOLBoundVar{x}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{C}\sp{\prime}}.
               \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{C}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
               \ensuremath{(}\ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                   \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
                   \ensuremath{[}\HOLBoundVar{E} \HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C} \HOLTokenTransBegin\HOLBoundVar{x}\HOLTokenTransEnd \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}}.
       \HOLConst{CCS_equation} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{Ys}. \HOLConst{MAP} \ensuremath{[}\HOLBoundVar{Ys}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{Es}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLSymConst{\ensuremath{=}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{Ys}. \ensuremath{[}\HOLBoundVar{Ys}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{n} \HOLBoundVar{xs} \HOLBoundVar{Ps} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
           \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_PROC} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}\ensuremath{)} \HOLBoundVar{Ps} \HOLSymConst{\HOLTokenConj{}}
           \HOLConst{TRACE} \ensuremath{(}\ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}\ensuremath{)} \HOLBoundVar{Ps}\ensuremath{)} \HOLBoundVar{xs} \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLBoundVar{xs} \HOLSymConst{\HOLTokenLeq{}} \HOLBoundVar{n} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{C}\sp{\prime\prime}}.
               \HOLConst{context} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{C}\sp{\prime\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \ensuremath{\HOLBoundVar{C}\sp{\prime\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
               \ensuremath{(}\ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime\prime}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}.
                   \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
                   \HOLConst{TRACE} \ensuremath{(}\ensuremath{(}\ensuremath{\HOLBoundVar{C}\sb{\mathrm{0}}} \HOLConst{\HOLTokenCompose} \HOLConst{FUNPOW} \HOLBoundVar{E} \HOLBoundVar{n}\ensuremath{)} \HOLBoundVar{Qs}\ensuremath{)} \HOLBoundVar{xs} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \ensuremath{\HOLBoundVar{C}\sp{\prime\prime}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}{\UseVerbatim{HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}
\HOLTokenTurnstile{} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{Y} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E} \HOLConst{DELETE} \HOLBoundVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXcombin}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLBoundVar{C}.
       \HOLConst{ALL_DISTINCT} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{Es} \HOLSymConst{\HOLTokenConj{}}
       \ensuremath{(}\HOLConst{LENGTH} \HOLBoundVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLBoundVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{[}\HOLBoundVar{Es}\ensuremath{/}\HOLBoundVar{Xs}\ensuremath{]} \HOLBoundVar{C}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXcombin}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXcombin}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXnil}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLConst{\ensuremath{\mathbf{0}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXnil}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXnil}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXpar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXpar}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXpar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXparXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXprefix}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXprefix}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXprefix}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrec}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y} \HOLBoundVar{E}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLBoundVar{Y} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLBoundVar{E} \HOLConst{DELETE} \HOLBoundVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLBoundVar{Xs}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrec}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrec}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelab}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelab}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelab}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestr}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestr}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestr}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXrules}
\HOLTokenTurnstile{} \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{u} \HOLBoundVar{E}. \HOLConst{context} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLBoundVar{u}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
        \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{\mid}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{L} \HOLBoundVar{E}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{nu} \HOLBoundVar{L} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{E} \HOLBoundVar{rf}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{relab} \HOLBoundVar{E} \HOLBoundVar{rf}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXrules}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXrules}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsum}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsum}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsum}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenEquiv{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}.
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}}
       \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXvar}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXvar}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXvar}}
\begin{SaveVerbatim}{HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Xs} \HOLBoundVar{Y}. \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLBoundVar{Y} \HOLBoundVar{Xs} \HOLSymConst{\HOLTokenImp{}} \HOLConst{weakly_guarded} \HOLBoundVar{Xs} \ensuremath{(}\HOLConst{var} \HOLBoundVar{Y}\ensuremath{)}
\end{SaveVerbatim}
\newcommand{\HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}{\UseVerbatim{HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule}}
\newcommand{\HOLMultivariateTheorems}{
\HOLThmTag{Multivariate}{BV_SUBSET_BIGUNION}\HOLMultivariateTheoremsBVXXSUBSETXXBIGUNION
\HOLThmTag{Multivariate}{CCS_solution_length}\HOLMultivariateTheoremsCCSXXsolutionXXlength
\HOLThmTag{Multivariate}{CCS_SUBST_elim}\HOLMultivariateTheoremsCCSXXSUBSTXXelim
\HOLThmTag{Multivariate}{CCS_SUBST_elim'}\HOLMultivariateTheoremsCCSXXSUBSTXXelimYY
\HOLThmTag{Multivariate}{CCS_SUBST_FEMPTY}\HOLMultivariateTheoremsCCSXXSUBSTXXFEMPTY
\HOLThmTag{Multivariate}{CCS_SUBST_FOLDR}\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDR
\HOLThmTag{Multivariate}{CCS_SUBST_FOLDR'}\HOLMultivariateTheoremsCCSXXSUBSTXXFOLDRYY
\HOLThmTag{Multivariate}{CCS_SUBST_IS_PROC}\HOLMultivariateTheoremsCCSXXSUBSTXXISXXPROC
\HOLThmTag{Multivariate}{CCS_SUBST_nested}\HOLMultivariateTheoremsCCSXXSUBSTXXnested
\HOLThmTag{Multivariate}{CCS_SUBST_nil}\HOLMultivariateTheoremsCCSXXSUBSTXXnil
\HOLThmTag{Multivariate}{CCS_SUBST_par}\HOLMultivariateTheoremsCCSXXSUBSTXXpar
\HOLThmTag{Multivariate}{CCS_SUBST_prefix}\HOLMultivariateTheoremsCCSXXSUBSTXXprefix
\HOLThmTag{Multivariate}{CCS_SUBST_PROC}\HOLMultivariateTheoremsCCSXXSUBSTXXPROC
\HOLThmTag{Multivariate}{CCS_SUBST_rec}\HOLMultivariateTheoremsCCSXXSUBSTXXrec
\HOLThmTag{Multivariate}{CCS_SUBST_reduce}\HOLMultivariateTheoremsCCSXXSUBSTXXreduce
\HOLThmTag{Multivariate}{CCS_SUBST_relab}\HOLMultivariateTheoremsCCSXXSUBSTXXrelab
\HOLThmTag{Multivariate}{CCS_SUBST_restr}\HOLMultivariateTheoremsCCSXXSUBSTXXrestr
\HOLThmTag{Multivariate}{CCS_SUBST_self}\HOLMultivariateTheoremsCCSXXSUBSTXXself
\HOLThmTag{Multivariate}{CCS_SUBST_sing}\HOLMultivariateTheoremsCCSXXSUBSTXXsing
\HOLThmTag{Multivariate}{CCS_SUBST_SING}\HOLMultivariateTheoremsCCSXXSUBSTXXSING
\HOLThmTag{Multivariate}{CCS_SUBST_sum}\HOLMultivariateTheoremsCCSXXSUBSTXXsum
\HOLThmTag{Multivariate}{CCS_SUBST_var}\HOLMultivariateTheoremsCCSXXSUBSTXXvar
\HOLThmTag{Multivariate}{context_backward_rules}\HOLMultivariateTheoremscontextXXbackwardXXrules
\HOLThmTag{Multivariate}{context_combin}\HOLMultivariateTheoremscontextXXcombin
\HOLThmTag{Multivariate}{context_nil}\HOLMultivariateTheoremscontextXXnil
\HOLThmTag{Multivariate}{context_par}\HOLMultivariateTheoremscontextXXpar
\HOLThmTag{Multivariate}{context_par_rewrite}\HOLMultivariateTheoremscontextXXparXXrewrite
\HOLThmTag{Multivariate}{context_par_rule}\HOLMultivariateTheoremscontextXXparXXrule
\HOLThmTag{Multivariate}{context_prefix}\HOLMultivariateTheoremscontextXXprefix
\HOLThmTag{Multivariate}{context_prefix_rewrite}\HOLMultivariateTheoremscontextXXprefixXXrewrite
\HOLThmTag{Multivariate}{context_prefix_rule}\HOLMultivariateTheoremscontextXXprefixXXrule
\HOLThmTag{Multivariate}{context_rec}\HOLMultivariateTheoremscontextXXrec
\HOLThmTag{Multivariate}{context_relab}\HOLMultivariateTheoremscontextXXrelab
\HOLThmTag{Multivariate}{context_relab_rewrite}\HOLMultivariateTheoremscontextXXrelabXXrewrite
\HOLThmTag{Multivariate}{context_relab_rule}\HOLMultivariateTheoremscontextXXrelabXXrule
\HOLThmTag{Multivariate}{context_restr}\HOLMultivariateTheoremscontextXXrestr
\HOLThmTag{Multivariate}{context_restr_rewrite}\HOLMultivariateTheoremscontextXXrestrXXrewrite
\HOLThmTag{Multivariate}{context_restr_rule}\HOLMultivariateTheoremscontextXXrestrXXrule
\HOLThmTag{Multivariate}{context_rules}\HOLMultivariateTheoremscontextXXrules
\HOLThmTag{Multivariate}{context_sum}\HOLMultivariateTheoremscontextXXsum
\HOLThmTag{Multivariate}{context_sum_rewrite}\HOLMultivariateTheoremscontextXXsumXXrewrite
\HOLThmTag{Multivariate}{context_sum_rule}\HOLMultivariateTheoremscontextXXsumXXrule
\HOLThmTag{Multivariate}{context_var}\HOLMultivariateTheoremscontextXXvar
\HOLThmTag{Multivariate}{DISJOINT_BV_CCS_SUBST}\HOLMultivariateTheoremsDISJOINTXXBVXXCCSXXSUBST
\HOLThmTag{Multivariate}{disjoint_imp_context}\HOLMultivariateTheoremsdisjointXXimpXXcontext
\HOLThmTag{Multivariate}{disjoint_imp_weakly_guarded}\HOLMultivariateTheoremsdisjointXXimpXXweaklyXXguarded
\HOLThmTag{Multivariate}{EVERY_weakly_guarded}\HOLMultivariateTheoremsEVERYXXweaklyXXguarded
\HOLThmTag{Multivariate}{FDOM_fromList}\HOLMultivariateTheoremsFDOMXXfromList
\HOLThmTag{Multivariate}{fromList_DOMSUB_NOT_IN_DOM}\HOLMultivariateTheoremsfromListXXDOMSUBXXNOTXXINXXDOM
\HOLThmTag{Multivariate}{fromList_EMPTY}\HOLMultivariateTheoremsfromListXXEMPTY
\HOLThmTag{Multivariate}{fromList_FAPPLY_EL}\HOLMultivariateTheoremsfromListXXFAPPLYXXEL
\HOLThmTag{Multivariate}{fromList_FAPPLY_EL'}\HOLMultivariateTheoremsfromListXXFAPPLYXXELYY
\HOLThmTag{Multivariate}{fromList_FAPPLY_HD}\HOLMultivariateTheoremsfromListXXFAPPLYXXHD
\HOLThmTag{Multivariate}{fromList_HD}\HOLMultivariateTheoremsfromListXXHD
\HOLThmTag{Multivariate}{FV_SUBSET_BIGUNION}\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNION
\HOLThmTag{Multivariate}{FV_SUBSET_BIGUNION_PRO}\HOLMultivariateTheoremsFVXXSUBSETXXBIGUNIONXXPRO
\HOLThmTag{Multivariate}{LIST_REL_equivalence}\HOLMultivariateTheoremsLISTXXRELXXequivalence
\HOLThmTag{Multivariate}{OBS_CONGR_subst_context}\HOLMultivariateTheoremsOBSXXCONGRXXsubstXXcontext
\HOLThmTag{Multivariate}{OBS_contracts_reflexive}\HOLMultivariateTheoremsOBSXXcontractsXXreflexive
\HOLThmTag{Multivariate}{OBS_contracts_subst_context}\HOLMultivariateTheoremsOBSXXcontractsXXsubstXXcontext
\HOLThmTag{Multivariate}{OBS_contracts_transitive}\HOLMultivariateTheoremsOBSXXcontractsXXtransitive
\HOLThmTag{Multivariate}{STRONG_EQUIV_subst_context}\HOLMultivariateTheoremsSTRONGXXEQUIVXXsubstXXcontext
\HOLThmTag{Multivariate}{strong_unique_solution_lemma}\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXlemma
\HOLThmTag{Multivariate}{strong_unique_solution_thm}\HOLMultivariateTheoremsstrongXXuniqueXXsolutionXXthm
\HOLThmTag{Multivariate}{unique_solution_of_obs_contractions}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXobsXXcontractions
\HOLThmTag{Multivariate}{unique_solution_of_rooted_contractions}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractions
\HOLThmTag{Multivariate}{unique_solution_of_rooted_contractions_lemma}\HOLMultivariateTheoremsuniqueXXsolutionXXofXXrootedXXcontractionsXXlemma
\HOLThmTag{Multivariate}{USC_unfolding_lemma1}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaOne
\HOLThmTag{Multivariate}{USC_unfolding_lemma2}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaTwo
\HOLThmTag{Multivariate}{USC_unfolding_lemma3}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaThree
\HOLThmTag{Multivariate}{USC_unfolding_lemma4}\HOLMultivariateTheoremsUSCXXunfoldingXXlemmaFour
\HOLThmTag{Multivariate}{weakly_guarded_backward_rules}\HOLMultivariateTheoremsweaklyXXguardedXXbackwardXXrules
\HOLThmTag{Multivariate}{weakly_guarded_combin}\HOLMultivariateTheoremsweaklyXXguardedXXcombin
\HOLThmTag{Multivariate}{weakly_guarded_imp_context}\HOLMultivariateTheoremsweaklyXXguardedXXimpXXcontext
\HOLThmTag{Multivariate}{weakly_guarded_nil}\HOLMultivariateTheoremsweaklyXXguardedXXnil
\HOLThmTag{Multivariate}{weakly_guarded_par}\HOLMultivariateTheoremsweaklyXXguardedXXpar
\HOLThmTag{Multivariate}{weakly_guarded_par_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXparXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_par_rule}\HOLMultivariateTheoremsweaklyXXguardedXXparXXrule
\HOLThmTag{Multivariate}{weakly_guarded_prefix}\HOLMultivariateTheoremsweaklyXXguardedXXprefix
\HOLThmTag{Multivariate}{weakly_guarded_prefix_rule}\HOLMultivariateTheoremsweaklyXXguardedXXprefixXXrule
\HOLThmTag{Multivariate}{weakly_guarded_rec}\HOLMultivariateTheoremsweaklyXXguardedXXrec
\HOLThmTag{Multivariate}{weakly_guarded_relab}\HOLMultivariateTheoremsweaklyXXguardedXXrelab
\HOLThmTag{Multivariate}{weakly_guarded_relab_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_relab_rule}\HOLMultivariateTheoremsweaklyXXguardedXXrelabXXrule
\HOLThmTag{Multivariate}{weakly_guarded_restr}\HOLMultivariateTheoremsweaklyXXguardedXXrestr
\HOLThmTag{Multivariate}{weakly_guarded_restr_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_restr_rule}\HOLMultivariateTheoremsweaklyXXguardedXXrestrXXrule
\HOLThmTag{Multivariate}{weakly_guarded_rules}\HOLMultivariateTheoremsweaklyXXguardedXXrules
\HOLThmTag{Multivariate}{weakly_guarded_sum}\HOLMultivariateTheoremsweaklyXXguardedXXsum
\HOLThmTag{Multivariate}{weakly_guarded_sum_rewrite}\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrewrite
\HOLThmTag{Multivariate}{weakly_guarded_sum_rule}\HOLMultivariateTheoremsweaklyXXguardedXXsumXXrule
\HOLThmTag{Multivariate}{weakly_guarded_var}\HOLMultivariateTheoremsweaklyXXguardedXXvar
\HOLThmTag{Multivariate}{weakly_guarded_var_rule}\HOLMultivariateTheoremsweaklyXXguardedXXvarXXrule
}