### 2019-12-15 ic-journal/intro.tex

\section{Introduction}

A prominent proof method for bisimulation, put forward by Robin Milner and widely used in his
landmark CCS book~\cite{Mil89}, is the
\emph{unique solution of equations}, whereby two tuples of processes are
componentwise bisimilar if they are solutions of the same system of equations.
This method is important in verification techniques and tools
based on algebraic reasoning~\cite{BaeBOOK,theoryAndPractice,RosUnder10}.

Milner's unique-solution theorem for weak bisimilarity, however,
has severe syntactic limitations:
the equations must be both \emph{guarded} and \emph{sequential}. That is,
the variables of the equations can only occur underneath visible prefixes and summation.
One way to overcome these limitations is to replace the equations
with special inequations called
\emph{contractions}~\cite{sangiorgi2015equations,sangiorgi2017equations}. The
contraction relation is a
preorder that, roughly, places some efficiency constraints on processes.
The unique solution of contractions is defined as with equations:
any two solutions must be componentwise bisimilar.
The difference from equations is in the meaning of a solution:
in the case of contractions the solution is evaluated with respect to
the contraction preorder rather than bisimilarity. 
With contractions, most syntactic limitations of the unique-solution theorem are
eliminated. One constraint that still remains
in~\cite{sangiorgi2017equations}
(where the issue is bypassed by using a more
restrictive CCS syntax)
is the occurrences of arbitrary sums (e.g.~$P + Q$) due to
the non-substitutivity of the contraction preorder in this case.

This paper presents a comprehensive formalisation of Milner's Calculus of Communicating
Systems (CCS) in the HOL theorem prover (HOL4),
with a focus towards the theory of unique solutions of equations and contractions.
Many results in Milner's CCS book~\cite{Mil89} are covered, since
the unique-solution theorems rely on a large number of fundamental results.
Indeed the formalisation encompasses all basic properties of strong,
weak and rooted
bisimilarities (e.g.~the fixed-point and substitutivity properties), and their algebraic laws.
Further extensions include several versions of ``bisimulation up to''
techniques, and properties of the expansion and contraction preorders.
Concerning rooted bisimilarity, the formalisation
includes Hennessy's Lemma, Deng's Lemma,
 and two theorems saying that rooted bisimilarity is the coarsest (largest)
 congruence contained in weak bisimilarity ($\wb$): one is classical
 with the hypothesis that no process uses up all labels;
the other one is 
without such hypothesis, essentially formalising van Glabbeek's proof~\cite{van2005characterisation}.
With this respect, the work is also an extensive experiment using the HOL theorem prover with its
recent developments, including its coinduction package.

This formalisation has offered us the possibility of
further refining the theory of unique solutions of contractions.
In particular,  existing results~\cite{sangiorgi2017equations} have
limitations on the body of the contractions due to the
substitutivity problems of weak bisimilarity and other behavioural relations with respect
to the sum operator.
In this paper, the contraction-based proof technique is further
refined by moving from the contraction preorder to
\emph{rooted contraction}, which is shown to be the coarsest
precongruence contained in the contraction preorder.
The resulting unique-solution theorem is now valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself), and places no 
constraints on summation.

Another benefit of the formalisation is 
that we can take advantage of results about different 
equivalences and preorders that share similar proof structures.
Such structural simililarities can be found, for instance, in the
following cases: the proofs that rooted bisimilarity and rooted
contraction are, respectively, the coarsest congruence contained in
weak bisimilarity and the coarsest precongruence contained in the
contraction preorder; the proofs about unique solution(s) of equations
for weak bisimilarity that use the contraction preorder as an
auxiliary relation, and other unique-solution theorems (e.g. the one
for rooted bisimilarity 
 in which the auxiliary relation is rooted contraction); the
proofs about various forms of enhancements of the bisimulation proof
method (the ``up-to'' techniques).
%
In these cases, when moving between proofs there are only a few places in
the HOL proof scripts that have to be modified.
Then the successful termination of a proof gives us the
 guarantee that the proof is correct, eliminating the risks 
of overlooking or missing details as in \emph{paper-and-pencil} proofs.

Concerning the formalisation of unique-solution theorems,
we first consider the case of a single equation (or contraction), the
\emph{\univariate case}.
Then we turn to the \emph{\multivariate case} where more equations
  (or contractions) with multiple equation variables are involved.
  The \univariate versions of the unique-solution theorems
  can directly use
$\lambda$-functions to represent CCS equations,
while the \multivariate versions require
more careful and delicate treatments of CCS
expressions, with multiple variables and substitutions acting on them.
% should be in the same paragraph, still talking about multivariable equations.
In contrast to literature such as~\cite{Gorrieri:2015jt},
we have followed Milner's original approach~\cite{milner1990operational} and adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
process is a CCS expression without equation variables.  % don't use "term"
This allows us a smoother move from the \univariate case to the \multivariate
one. (See Section~\ref{sec:multivariate} for more details.)

\paragraph{Structure of the paper}

In Section~\ref{ss:ccs} we recall the core theory of CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Then, Section~\ref{s:eq} discusses equations and contractions, and in particular,
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
In Section~\ref{s:for} we highlight the CCS formalisation in HOL4, with
the unique-solution theorems in the \univariate case.
Section~\ref{sec:multivariate} describes the extension to the \multivariate case.
Finally, in Section~\ref{s:rel} and \ref{s:concl} we discuss the related work,
conclusions, and a few directions for future work.

%%% next file: ccs.tex